Mega-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-11-04 paper/1_Initial/manuscript.qmd
minor

<!-- Mood and psychopathology -->

@@

Among Big Five traits, Openness (mean β = .10), Agreeableness (β = .08), and Conscientiousness (β = .11) showed small but consistent positive effects, most evident for the full IAS. Honesty–Humility (mean β = −.10) showed a small negative effect, primarily for Hungry–Thirsty and Bruise–Blood sugar. Neuroticism (mean β = −.13) was negatively associated with the IAS, with the strongest β for Hungry-Thirsty (β = −.23).

@@

<!-- Interoception and correlates - item specific -->
The IAS showed positive associations with most MAIA facets, except for the Not-Distracting and Not-Worrying subscales, replicating prior findings [@brand2023]. Associations with the BPQ Body Awareness subscale were consistent with previous reports on positive association between these measures [@brand2023; @campos2021; @koike2023; @gaggero2021], whereas indices of autonomic reactivity were more strongly and negatively associated with IAS scores, supporting earlier observations that reactivity may reflect aspects of bodily signal processing captured by self-reported interoceptive accuracy [@todd2022].

@@

Across mood and interoceptive variables, most item pairings displayed small-to-moderate standardized coefficients, indicating shared variance between interoceptive self-reports and affective tendencies. In contrast, associations with psychopathology were more selective: negative associations were primarily observed for visceral pairings such as Hungry–Thirsty and Bruise–Blood sugar, whereas other domains yielded weaker or nonsignificant effects. This pattern suggests that self-reported interoceptive accuracy is more strongly associated with affective and clinical traits for visceral sensations than for somatic or respiratory reflexes [@murphy2019; @koike2023].

@@

Associations with personality traits followed a similar domain-specific structure. Effects for the full IAS largely reflected contributions from specific pairings: Extraversion showed its strongest associations for Breathing–Heart, Neuroticism and Honesty–Humility for Hungry–Thirsty, and Openess, Conscientiousness and Agreeableness for the full scale. These results indicate that associations between interoceptive self-report and personality dimensions may depend on the bodily domain represented.

@@

<!-- Short IAS benefits and summary-->
The 14-item IAS performed similarly to the original 21-item version, showing comparable patterns of association with mood, psychopathology, and interoceptive measures. Overall, standardized effects between the IAS and external constructs were not uniform but varied across bodily domains. Visceral item pairs, particularly Hungry–Thirsty and Bruise–Blood sugar, showed the strongest and most consistent relationships with interoceptive, affective, and personality measures, whereas other pairs contributed little unique information, particularily in measures of psychopathologies and beliefs. These findings suggest that the shorter scale efficiently preserves these domain-specific associations, offering a more parsimonious assessment of self-reported interoceptive accuracy, though further validation across independent samples is warranted.

>>>>

<!-- Mood-->

@@

Among the Big Five traits, Openness (mean β = .10), Agreeableness (mean β = .08), and Conscientiousness (mean β = .11) exhibited small yet consistent positive effects, most evident for the full IAS. Extraversion showed a weak but positive association, driven primarily by its stronger relation to the Bruise–Blood pairing (β = .17). In contrast, Honesty–Humility (mean β = −.10) demonstrated a small negative effect, most evident for the Hungry–Thirsty and Bruise–Blood sugar pairings. Finally, Neuroticism (mean β = −.23) was negatively associated with the IAS, with the strongest effect observed for the Hungry–Thirsty pairing (β = −.22).

@@

<!-- Interoception and correlates - item specific -->
The IAS showed positive associations with most MAIA facets, except for the Not-Distracting and Not-Worrying subscales, replicating prior findings [@brand2023; @todd2022]. Associations with the BPQ Body Awareness subscale were consistent with previous reports on positive association between these measures [@brand2023; @campos2021; @koike2023; @gaggero2021], whereas indices of autonomic reactivity were more strongly and negatively associated with IAS scores, supporting earlier observations that reactivity may reflect aspects of bodily signal processing captured by self-reported interoceptive accuracy [@todd2022].

@@

Across mood, most item pairings displayed small-to-moderate standardized coefficients, indicating shared variance between interoceptive self-reports and affective tendencies. In contrast, associations with psychopathology were more selective: negative associations were primarily observed for visceral pairings such as Hungry–Thirsty and Bruise–Blood sugar, whereas other domains yielded weaker or nonsignificant effects. This pattern suggests that self-reported interoceptive accuracy is more strongly associated with affective and clinical traits for visceral sensations than for somatic or respiratory reflexes [@murphy2019; @koike2023].

@@

Associations with personality traits followed a similar domain-specific structure. Effects for the full IAS largely reflected contributions from specific pairings: Extraversion showed its strongest associations for Bruise-Blood sugar, Neuroticism and Honesty–Humility for Hungry–Thirsty, and Openess, Conscientiousness and Agreeableness for the full scale. These results indicate that associations between interoceptive self-report and personality dimensions may depend on the bodily domain represented.

@@

<!-- Short IAS benefits and summary-->
The 14-item IAS performed similarly to the original 21-item version, showing comparable patterns of association with mood, psychopathology, and interoceptive measures. Overall, standardized effects between the IAS and external constructs were not uniform but varied across bodily domains. Visceral item pairs, particularly Hungry–Thirsty and Bruise–Blood sugar, showed the strongest and most consistent relationships with interoceptive, affective, and personality measures, whereas other pairs contributed little unique information, particularily in measures of psychopathologies and beliefs. These findings suggest that the shorter scale captures domain-specific patterns that were not evident in the full scale, offering a more efficient and nuanced measure of self-reported interoceptive accuracy, though further validation in independent samples is needed.
